---
title: "Pima Indians Diabete"
author: "Mehedi Hasan Mir"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
    code_folding: show
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE, message = FALSE)
```

# Packages

Loading required packages:


```{r}
library(dplyr)
library(knitr)
library(DescTools)
library(ggplot2)
library(naniar)
library(tidyr)
```


------------------------------------------------------------------------

# Importing the Pima Indians Diabetes Data


```{r}
df = read.csv("Data/Kaggle - Pima Indians Diabetes Database.csv")
head(df)
```


Here is what the columns in the data mean:

| Column | Details |
|--|----|
| Pregnancies | Number of times pregnant |
| Glucose | Plasma glucose concentration a 2 hours in an oral glucose tolerance test |
| BloodPressure | Diastolic blood pressure (mm Hg) |
| SkinThickness | Triceps skin fold thickness (mm) |
| Insulin | 2-Hour serum insulin (mu U/ml) |
| BMI | Body mass index (weight in kg/(height in m)\^2) |
| DiabetesPedigreeFunction | Diabetes pedigree function |
| Age | Age (years) |
| Outcome | Class variable (0=No or 1=Yes) |

------------------------------------------------------------------------


**`Outcome` is a categorical variable, 1 means Yes, 0 means No. So, convert it into a factor data type. Then, calculating mean, standard deviation, minimum, 1st quartile, median, 3rd quartile, maximum, and inter-quartile range of all the numeric columns in the dataset.
Here, In coloumn of Glucose, BloodPressure, SkinThickness & BMI, There are lots of 0 that is 
meaningless. In real life O in not exist in these columns.

```{r}
df$Outcome = as.factor(df$Outcome)
class(df$Outcome)
calculating = function(colum){
  colum = as.numeric(colum)
  mean_value = mean(colum)
  median_value = median(colum)
  mode_value = Mode(colum)
  min_value = min(colum)
  standard_value = sd(colum)
  max_value = max(colum)
  quantile = quantile(colum)
  first_quantile = quantile(colum, 0.25)
  third_quantile = quantile(colum, 0.75)
  iqr_range = IQR(colum)
  
  return(list(
    mean = mean_value,
    median = median_value,
    mode= mode_value,
    min = min_value,
    max = max_value,
    standard_deviation = standard_value,
    quantile_ = quantile,
    firstQuantile = first_quantile,
    thirdQuantile = third_quantile,
    iqrRange = iqr_range
  ))
}
calculating(df$Pregnancies)
calculating(df$Glucose)
calculating(df$BloodPressure)
calculating(df$SkinThickness)
calculating(df$Insulin)
calculating(df$BMI)
calculating(df$DiabetesPedigreeFunction)
calculating(df$Age)
calculating(df$Outcome)
```


**Let's see how many people have Outcome = Yes?**


```{r}
yes_outcone = df %>% 
  filter(Outcome == 1)
count(yes_outcone)
```


**Creating histograms for all the numeric variables in the data using ggplot2.**


```{r}
histplot = function(colum){
  colum = as.numeric(colum)
  ggplot(data = df) +
    geom_histogram(aes(x = colum, fill = Outcome, colour = "black", alpha = 0.8),
                   bins = 20)
}
histplot(df$Pregnancies)
```
```{r}
histplot(df$Glucose)
```
```{r}
histplot(df$BloodPressure)
```
```{r}
histplot(df$SkinThickness)
```
```{r}
histplot(df$Insulin)
```
```{r}
histplot(df$BMI)
```
```{r}
histplot(df$DiabetesPedigreeFunction)
```
```{r}
histplot(df$Age)
```



**Creating boxplot of Glucose, Insulin, BMI, and Age by Outcome using ggplot2.**


```{r}
df %>% 
  ggplot(aes(x = Outcome, y = Glucose))+
   geom_boxplot()+
  geom_jitter()

```
```{r}
df %>% 
  ggplot(aes(x = Outcome, y = Insulin))+
  geom_boxplot()+
  geom_jitter()
  
```
```{r}
df %>% 
  ggplot(aes(x = Outcome, y = BMI))+
  geom_boxplot()+
  geom_jitter()
```
```{r}
df %>% 
  ggplot(aes(x = Outcome, y = Age))+
  geom_boxplot()+
  geom_jitter()
```


**Replacing 0 with NA in the variables where a value of 0 does not make sense.**


```{r}
null_column = colSums(is.na(df))
zero_col = colSums(df==0)
df$Glucose[df$Glucose == 0] = NA
df$BloodPressure[df$BloodPressure == 0] = NA
df$SkinThickness[df$SkinThickness == 0] = NA
df$BMI[df$BMI == 0] = NA
head(df)
```



**Useing naniar package to inspect number of missing values in the data after replacing 0's with NA's**


```{r}
miss_var_summary(df)
gg_miss_var(df)
vis_miss(df)
```
```{r}
summary(df)
nrow(df)
df1 =df %>% 
  mutate(SkinThickness = ifelse(is.na(SkinThickness), mean(SkinThickness, na.rm = TRUE), SkinThickness),
         BloodPressure = ifelse(is.na(BloodPressure), median(BloodPressure, na.rm = TRUE), BloodPressure))
final_df = df1 %>% 
  drop_na(Glucose, BMI)
miss_var_summary(final_df)
```


**Export the final cleaned data in an Excel file for modeling.**

```{r}
write.csv(final_df, "Output/Processed Pima Indians Diabetes.csv", row.names = F)
```


